---
title: "Problem Set 6"
subtitle: "Exploratory Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document: 
    highlight: textmate
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

The objective of this assignment is to familiarize your self with the process, specifically the visual process of Exploratory Data Analysis (EDA).  I have assigned several readings to support this work and layout an overall procedure.  You should read **Zuur et al.** prior to completing this work.


## Directions

Using any data set in `R`, for instance the `titatic` data set found within the `library(datasets)` package as your analysis data.  Conduct an EDA on the data set of your choice.  Your analysis should follow the "key steps" discussed in **Zuur et al.** and in our lecture discussion.  

You analysis should be completed within an **Rmarkdown** document.  The assignment should show data import, and any data cleanup that you conduct.  The document should show the steps in your EDA including:

- $\geq$ 5 statistical graphics.  
- Each graphic should include plot titles, proper axis labels, and legend information, if neccesary.  
- Along with each plot include a short statement indicating what you learned from that plot.  (This could be anything from verifying a `sort()` or `filter()` call to a `mosaic()` plot showing sample size variability).  

The final document should be converted using the `knitr` package to an `.html` file and submitted via Moodle for evaluation.  Remember to properly name your file as: `YourName_ANLY512-50_Fall.html`.

### Calling all the helpers

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(magrittr)
library(datasets)

```

For this analysis, I am going to explore the diamonds dataset with the ggplot 2 packages, which could be summarised by `glimpse()` 

```{r}

glimpse(diamonds)

```

### Plot 1: Dimond Carat Distribution

```{r}

diamonds %>% 
    ggplot(aes(carat, fill = factor(cut))) + 
    geom_density(alpha = 0.5) + 
    ggtitle("Diamonds Carat Distribution by Cut") +
    labs(x="Carat",y="Density") 

```

> There are clearly clusters in the carat distribution and most diamonds are under 3 carats


### Plot 2: Relationship between Diamonds Carat and Price

```{r}
diamonds %>% 
    filter(carat < 3) %>% 
    ggplot(aes(x=carat, y=price)) + 
    geom_point(alpha = 0.5, size = 2, color = "#004D43") + 
    stat_smooth(method="lm") + 
    ggtitle("Relationship between diamonds carat and price") +
    labs(x="Carat",y="Price")
```

> There is a significant positive relationship btw diamonds carat and its price. 

### Plot 3: Relationship between diamonds price and cut. 

```{r}

diamonds %>% 
    ggplot(aes(cut, price)) + 
    geom_boxplot(outlier.shape = NA, width  = 0.5, alpha = 0.7, color="#4C4C4C", notch = T) + 
    coord_flip() + 
    geom_point(color = "#004D43", alpha = 0.2, size = 1) + 
    ggtitle("Relationship between diamonds price and different cut") +
    labs(x="Price",y="Cut")

```

> Premium cut seems to have highest median price

### Plot 4: Relationship between cut and color 

```{r}
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n)) + 
    ggtitle("Relationship between diamonds color and different cut") +
    labs(x="Color",y="Cut")
```

> ideal cut and color G is most common

### Plot 5: Relationship between Diamond Depth and Price

```{r}
diamonds %>% 
    filter(carat < 3) %>% 
    ggplot(aes(x=factor(clarity), y=price)) + 
    geom_bin2d() +
    ggtitle("Relationship between diamonds clarity and price") +
    labs(x="Clarity",y="Price")
```

> Most common clarity combination is under $2000 with VS2 clarity


